---
title: "P8105 - Homework 3"
author: "Molly Martorella"
date: "10/12/2019"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(viridis)
library(p8105.datasets)

knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%")

options(knitr.kable.NA = "")

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_bw())

```

# Problem 1

The Instacart data from 2017 has **`r nrow(instacart)`** orders, and it describes these orders with **`r ncol(instacart)`** variables, some of which are: _order_id_, _product_id_, _product_name_, whether it was _reordered_, and the _department_id_ and _aisle_. There are 134 aisles, and the fresh vegetable and fresh fruit aisles are ordered from the most.

## Plot of Most Popular Aisles

```{r include = FALSE}

data("instacart")

instacart %>% 
  count(aisle)

# 134 aisles
```

```{r}

instacart %>% 
  count(aisle, sort = TRUE) %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = reorder(aisle, -n), y = n)) +
  geom_col() +
  ggtitle("Number of Instacart Orders by Aisle") +
  xlab("Aisle") +
  ylab("Number of Items Ordered") +
  theme(axis.text.x = element_text(angle = 60, 
                                   hjust = 1))

```

## Three most popular items from select aisles

The number of times the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” are ordered is presented in the following table:

```{r}

instacart %>% 
  filter(aisle %in% c("baking ingredients", 
                    "dog food care", 
                    "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name, sort = TRUE) %>% 
  filter(min_rank(desc(n)) <= 3) %>% 
  pivot_wider(names_from = aisle,
              values_from = n) %>% 
  knitr::kable()

```

## Ordering Day/Time of Pink Lady Apples vs Coffee Ice Cream

The following is a table showing the mean hour of the day per each day of the week that Pink Lady Apples and Coffee Ice Cream are ordered.

```{r}

instacart %>% 
  filter(product_name %in% c("Pink Lady Apples",
                           "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  mutate(mean_hour = mean(order_hour_of_day)) %>% 
  select(c(product_name, 
           order_dow, 
           mean_hour)) %>%
  distinct() %>% 
  pivot_wider(names_from = order_dow, 
              values_from = mean_hour) %>% 
  rename(Monday = '0',
         Tuesday = '1',
         Wednesday = '2',
         Thursday = '3',
         Friday = '4',
         Saturday = '5',
         Sunday = '6') %>%
  select(product_name, 
         Monday, 
         Tuesday, 
         Wednesday, 
         Thursday, 
         Friday, 
         Saturday, 
         Sunday) %>% 
  knitr::kable(digits = 1)

```






